\thetitle

### 2021-05-17 paper/paper.tex
litrev

\begin{abstract}
  % teilw. von IJPR paper uebernommen
  An essential task in manufacturing planning and control is to determine when to release orders to
  the shop floor. One key parameter is the lead time which is the planned time that elapses between
  the release of an order and its completion. Lead times are usually determined based on the
  measured duration orders take to traverse the production system, i.e.\@ the flow times.
  Traditional order release models assume static lead times, although it has been shown that they
  should be set dynamically to be able to react to the dynamic operational characteristics of
  production system. Therefore, we present an order release model which sets lead times dynamically
  by using an reinforcement learning approach. This algorithm is designed especially for
  applications with high process variability and periodic feedback to the agent. Therefore, this
  paper aims at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our
  method on a three-stage flow-shop simulation model not only to static order release techniques,
  but also to the commonly used deep Q-Learning algorithm. The results show that both reinforcement
  learning agents have room for improvement for queuing systems with low variability, that our
  proposed method usually outperforms deep Q-Learning and finally for systems with high utilisation
  that the proposed reinforcement learning algorithm performs very well, outperforming all other
  approaches.
  %
  % We use a  to compare the algorithm
  % to % and show that our approach outperforms
  % well-known order release mechanisms. We show that in the current version our proposed model using
  % reinforcement learning outperforms some, but not all other tested approaches.
  % EVTL: especially for scenarios with...
  \keywords{operations research, production planning, order release, machine learning, reinforcement
    learning}
\end{abstract}

@@

Setting of lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
release mechanisms use static or fixed lead times to address the order release problem.
%
However, lead times should be set dynamically due to the non-linear relationship between utilisation
and flow times \citealt{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
system performance. For instance, a solely reactive approach can result in the lead time syndrome.
%
The \textit{lead time syndrome} describes a vicious cycle where increasing flow times perpetually
inflate the lead times which leads to worse performance (see e.g., \citealt{Mather1978,
  knollmann2013control, Selcuk2006}).
%
%
\citet{schneckenreither2020order} have established following categorisation of dynamic lead time
management approaches:

@@

\MS[]{todo: Reactive}

\MS[]{todo: proactive}

We hypothesise that dynamic lead time management approaches need to aim for a predictive design in
order to be able to compete with state-of-the-art order release methods from literature. However,
there only exist three papers that propose a predictive lead time management algorithm in
literature.

The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
using reinforcement learning in a single product, four-station serial flow line and compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
%
% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
% is to find the best stationary policy for a given problem. This policy is usually provided by
% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
% (MDP).
% The advantage of reinforcement learning over dynamic programming is that (i) the problem
% space is explored by an agent and thus only expectantly interesting parts of the problem space need
% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
% as the states are evaluated by consecutively observed states solely.
%
The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
each system change, the completion of an operation of any order at any stage or a new order arrival,
and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
although in practice order release decisions often need to be made on a periodical basis, e.g.\@
daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the syndrome. Their model outperforms static and reactive lead time
management approaches, especially for scenarios with high utilisation.

@@

For the experimental runs we used personal computers with Intel CPUs i7-7700 (16GB RAM) and i9-7900X
(32GB RAM). For each method we performed \(500.000\) learning steps, which took about \(2\)-\(3\)
days. For the same experimental setup the inputs/random variables are the same for each period.
After learning each order release method was evaluated on \(20\) demand streams, each with \(1000\)
periods of warm-up and \(7000\) periods of evaluation phase. Welch’s procedure was applied to
approximate the length of the warm-up period (see~\cite{law:simulationc}).

@@

\begin{algorithm}[h!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
    set exploration rate \(0 < \pexp \leqslant 1\), target network adaption rate
    \(0 < \gamma \ll 1\) and discount factor \(0 < \gamma_{1} < 1\)
    \While{the stopping criterion is not fulfilled} \ForEach{independent agent
      \(i \in \agent_{1},\agent_{2},\ldots,\agent_{n}\)} \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise one of
      the set defined by
      \(\min_{a \in \mathcal{A}(i)}\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{})\).
    \end{minipage}
    \EndFor{}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
      Carry out action \(a_{t}=(a_{t,\agent_{1}},\ldots,a_{t,\agent_{n}})\), observe reward
      \(r_{t}\) and resulting state \(s_{t+1}\). Store experience
      \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay memory \(M\).
    \end{minipage}
    \If{\(t \mod{\nstep} \equiv 0\)}
    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      Sample \(m\) random mini-batches of \(\nstep\) consecutive experiences from \(M\) numbered by
      \(1 \leqslant i \leqslant \nstep\), each with
      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
    \end{minipage}
    \ForEach{mini-batch}
    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
    \If{experience \(\nstep\) or \(\neg \isRandAct{i}\) or  \(\pexp > \plearn\)}
      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{i+1},a; \netTarget{})\)}
    \EndIf
    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} \)}
    \State{\(d\netWorker{} \gets d\netWorker{} +
      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \EndFor{}
    \EndFor{}
    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{} / (m \cdot \nstep)\)}
    \EndIf{}
    \State{Every \(C\) steps exponential smoothly set target network:
    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
    \EndWhile{}
  \end{algorithmic}
  \caption{\label{alg:ql}Deep Q-Learning}
\end{algorithm}

>>>>

\begin{abstract}
  % teilw. von IJPR paper uebernommen
  A major task in hierarchical structured manufacturing planning and control is to determine when to
  release orders to the shop floor. One key parameter is the lead time which is the planned time
  that elapses between the release of an order and its completion. Lead times are usually determined
  based on the flow times, i.e.\@ the measured duration orders take to traverse the production
  system. Traditional order release models assume static lead times, although it has been shown that
  they should be set dynamically to be able to react to the dynamic operational characteristics of
  production system. Therefore, we present a reinforcement learning order release model which sets
  lead times dynamically. The applied reinforcement learning agent is designed especially for
  processes with periodic feedback and high variability. This paper aims at providing a
  proof-of-concept for the proposed algorithm. We compare our method on a three-stage flow-shop
  simulation model not only to static order release techniques, but also to the commonly used deep
  Q-Learning agent. The results show that both reinforcement learning agents have room for
  improvement for queuing systems with low variability, that our proposed method usually outperforms
  deep Q-Learning and finally for systems with high utilisation that the proposed reinforcement
  learning algorithm performs very well, outperforming all other approaches.
  %
  % We use a  to compare the algorithm
  % to % and show that our approach outperforms
  % well-known order release mechanisms. We show that in the current version our proposed model using
  % reinforcement learning outperforms some, but not all other tested approaches.
  % EVTL: especially for scenarios with...
  \keywords{operations research, production planning, order release, machine learning, reinforcement
    learning}
\end{abstract}

@@

Setting lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
release mechanisms use static or fixed lead times to address the order release problem.
%
However, lead times should be set dynamically due to the non-linear relationship between utilisation
and flow times \citealt{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
system performance. For instance, a solely reactive approach can result in the lead time syndrome.
%
The \textit{lead time syndrome} describes a vicious cycle where increasing flow times perpetually
inflate the lead times which leads to worse performance (see e.g., \citealt{Mather1978,
  knollmann2013control, Selcuk2006}).
%
%
\citet{schneckenreither2020order} have established following categorisation of dynamic lead time
management approaches:

@@

% \MS[]{todo: Reactive}

In the sequel we review the lead time management (LTM) literature for due date assignment, order
release and scheduling, with the focus on order release literature.

\paragraph{Reactive LTM.}
%
%% Reactive Due Date Ass
Early due date assignment approaches \citep[e.g.][]{conway1967} set the due dates according to the
processing time (Total Work Content rule, TWK rule) or the number of operations.
%
%% Reactive Order Release
For order release \citet{enns2004work} investigate the use of a reactive approach with exponential
smoothing of flow times to set lead times for two-stage distribution requirements planning. Although
it can be considered an exploratory research paper the main conclusions are that dynamically setting
planned lead times using exponential smoothing works well, but that the reactions are delayed. The
later is due to the nature of the flow time measurement only being possible for finished orders.
%
Another reactive approach uses exponential smoothing to forecast flow times to set lead times for a
capacitated multi-stage make-to-stock system \citep{Selcuk2006}. Their approach performs better than
fixed lead time techniques for most experiments. However, they also exhibit the LTS\@.

\paragraph{Proactive LTM.}
%
%% Proactive Due Date Assignment
The TWK due date assignment rules have been extended by several researchers to proactive methods
\citep{eilon1976, kanet1986, Enns1995}. As flow times depend on the workload and machine capacity in
the shop workload dependent \citep{Bertrand1983} due date assignment rules have been developed. The
basic idea is to estimate the flow time of an order on its arrival by measuring the number of orders
queued on the corresponding routing, which can be analytically \citep[e.g.][]{ChungHuang2002}, by
the use of a simulation model \citep[e.g.][]{vig1991}, statistical methods \citep[e.g.][]{Tai2012}
or using hybrid approaches \citep[e.g.][]{Kaplan1993}. Furthermore several proactive due date
assignment methods that incorporate neural networks exist \citet{Philipoom1994, Philipoom1997,
  HsuSha2004,Patil2008,ChangetalFuzzyDDset2008}.
%
%% Proactive Order Release
In the case of order release the analytical papers by \citet{Selcuk2009} and \citet{Selcuk2013} use
the WIP level to quote lead times. In the former they show that the less frequently updating the
lead time decreases the nervousness and improves the performance, while in the latter that
dynamically updating lead times performs much better than static lead times.
%
% MIL (Modified infinite loading)
MIL extends the static BIL procedure to integrate shopfloor congestion by subtracting number of
queued jobs on the corresponding routing from the due date
\citep{RagatzMabert1988,wisner1995review}. For non-due-date oriented dispatching rules, their
approach has a positive effect on total cost performance, while for due-date oriented dispatching
this holds only if the due dates are loose. For all other cases the release algorithm has no
negative effect on performance. Variants of this rule-based method exist
\citep[e.g.][]{park1989job,kim1995evaluating}.
% Bertrand 2002
% \citet{bertrand2002investigation} investigate different ways of releasing orders, assigning lead
% times and operation due dates to assembly orders. In their work a due date is set once the order
% arrives and then a release time calculated by subtracting the lead time from it. In one approach the
% lead time calculation is the sum of the operation times plus a factor consisting of the waiting time
% allowance per operation times the number of operations. They use the average operation waiting time
% as waiting time allowance factor. In the other approaches the orders are released using the
% \ac{imre} policy. In terms of order release they find that all work orders of an assembly order
% should be released at the same time.
%
A machine learning approach is presented by \citet{Lee1997}. They combine a genetic algorithm with a
decision tree to generate sequencing rules in the order pool for release. In the tested job shops
the CONWIP rule determines the release of the orders and a decision tree to find the order release
sequence. Their method leads to significant improvements over conventional single release and
dispatching approaches.

%% Predictive
\paragraph{Predictive LTM.}
%
Due to the complex problem structure we hypothesise that dynamic lead time management approaches
should aim for a predictive design in order to be able to compete with state-of-the-art order
release methods from literature. However, there only exist four papers that propose a predictive
lead time management method.

For scheduling \citet{wang2007reinforcement} present a predictive lead time management approach
using reinforcement learning. Their algorithm uses multiple agents to forecast the processing time
of each job for routing decisions. They show on a 10 machine job shop that the RL algorithm performs
equally or better than tested heuristics.
%
In case of order release, \citet{PaterninaArboleda2001} introduce a predictive order release model
using reinforcement learning in a single product, four-station serial flow line. They compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP). The
algorithm decides on whether or not to release an order after each system change, the completion of
an operation of any order at any stage or a new order arrival, and assumes that any unsatisfied
demand is lost. Thus, they use a continuous order release method although in practice order release
decisions often need to be made on a periodical basis, e.g.\@ daily (see
\citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their
tabular based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the syndrome. Their model outperforms static and reactive lead time
management approaches, especially for scenarios with high utilisation.

@@

For the experimental runs we used personal computers with Intel CPUs i7-7700 (16GB RAM) and i9-7900X
(32GB RAM). For each method we performed \(500.000\) learning steps, which took about \(2\)-\(3\)
days. For the same experimental setup the inputs/random variables are the same for each period.
After learning, each order release method was evaluated on \(20\) demand streams, each with \(1000\)
periods of warm-up and \(7000\) periods of evaluation phase. Welch’s procedure was applied to
approximate the length of the warm-up period (see~\cite{law:simulationc}).

@@

\begin{algorithm}[h!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
    set exploration rate \(0 < \pexp \leqslant 1\), target network adaption rate
    \(0 < \gamma \ll 1\) and discount factor \(0 < \gamma_{1} < 1\)
    \While{the stopping criterion is not fulfilled} \ForEach{independent agent
      \(i \in \agent_{1},\agent_{2},\ldots,\agent_{n}\)} \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise the one
      fulfilling
      \(\min_{a \in \mathcal{A}(i)}\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{})\).
    \end{minipage}
    \EndFor{}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
      Carry out action \(a_{t}=(a_{t,\agent_{1}},\ldots,a_{t,\agent_{n}})\), observe reward
      \(r_{t}\) and resulting state \(s_{t+1}\). Store experience
      \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay memory \(M\).
    \end{minipage}
    \If{\(t \mod{\nstep} \equiv 0\)}
    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      Sample \(m\) random mini-batches of \(\nstep\) consecutive experiences from \(M\) numbered by
      \(1 \leqslant i \leqslant \nstep\), each with
      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
    \end{minipage}
    \ForEach{mini-batch}
    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
    \If{experience \(\nstep\) or \(\neg \isRandAct{i}\) or  \(\pexp > \plearn\)}
      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{i+1},a; \netTarget{})\)}
    \EndIf
    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} \)}
    \State{\(d\netWorker{} \gets d\netWorker{} +
      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \EndFor{}
    \EndFor{}
    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{} / (m \cdot \nstep)\)}
    \EndIf{}
    \State{Every \(C\) steps exponential smoothly set target network:
    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
    \EndWhile{}
  \end{algorithmic}
  \caption{\label{alg:ql}Deep Q-Learning}
\end{algorithm}
